Crop adaptation through gene flow

### 2016-06-28 Crop_adaptation_gene_flow_gj_outline_shortened.tex
Update Crop_adaptation_gene_flow_gj_outline_shortened.tex

\subsection*{Introduction}

 "Whenever sympatric populations of crops and their wild relatives show a similar trait, there will be several possible explanations: (1) retention of an ancestral trait, (2) convergent or parallel evolution, and (3) introgression.  Perhaps even more difficult is establishing the direction of introgression; does gene flow go from crop into wild species, the reverse, or in both directions?" and "For this reason, introgression may be best studied between crops and their more distant wild relatives where the probability of finding taxon-specific genetic markers is increased." \cite{doebley1990molecular}

@@

@@

Olive

Morphological similarities indicate that wild, weedy, and cultivated olives form an interfertile species complex that sprawls across the Mediterranean, and that outbreeding with these wild and weedy relatives has contributed to olive's genetic diversity, but it has not been shown that locally-adapted genes have been introgressed into the domesticated olive \cite{zohary1975beginnings}\.

\subsection*{Sorghum}
Anthropological evidence points to a domestication event of sorghum (\emph{Sorghum bicolor subsp. bicolor}) some 5-6,000 years ago in Ethiopian Sudan \cite{smith2000sorghum}\.
However, comparative isozyme analyses \cite{shechter1975comparative}\ have suggested a possible secondary domestication event (sorghum race kafir) in the southeastern Savanna.
The progenitor species was \emph{S. arundinaceum} \cite{doggett1988sorghum, harlan1971toward}\.
The genus Sorghum contains 25 species, but these are split between five subgenera: Eu-Sorghum, Chaetosorghum, Heterosorghum, Para-Sorghum, and Stiposorghum, and it is Eu-Sorghum that includes all domesticated, weedy, and closely-related wild relatives \cite{USDAARS2007, garber1950cytotaxonomic}.
Within Eu-Sorghum, there are currently nine races, five cultivated (bicolor, guinea, kafir, caudatum, and durra) and four wild (arundinaceum, virgatum, aethiopicum, and verticilliflorum) \cite{smith2000sorghum}\.
Members of \emph{Sorghum bicolor} are interfertile, though outcrossing rates and fertility can be low due to ploidy incompatability \cite{doggett1988sorghum, arriola1996crop}.

@@

/subsection*{Soybean}
Soybeans (Glycine max) were domesicated in southern China \cite{guo2010single}\.
Soybeans (Glycine max) primarily self-pollinate, outcrossing at a rate of 1-8% \cite{ray2003soybean}\.
Soybeans readily cross with their wild progenitor, G. soja \cite{singh1988genomic}\.
Plants of intermediate morphology appear around fields when S. soja is also present, indicating spontaneous hybridization \cite{kwon1972studies}\.
The accession "G. gracilis" shows allelic contribution from both G. max and G. soja \cite{keim1989restriction}\.

>>>>

@@

\subsection*{Soybean}
Soybeans (Glycine max) were domesicated in southern China \cite{guo2010single}\.
Soybeans (Glycine max) primarily self-pollinate, outcrossing at a rate of 1-8% \cite{ray2003soybean}\.
Soybeans readily cross with their wild progenitor, G. soja \cite{singh1988genomic}\, to the degree that some researchers \cite{hymowitz1980taxonomy} consider the two to be conspecific.
Plants of intermediate morphology appear around fields when S. soja is also present, indicating spontaneous hybridization \cite{kwon1972studies}\.
The accession "G. gracilis" shows allelic contribution from both G. max and G. soja \cite{keim1989restriction}\.
Presently, however, there is little or no evidence that natural wild-to-crop introgressive hybridization has had an adaptive impact on cultivated soybeans.

\subsection*{Soybean}
Soybeans (Glycine max) were domesicated in southern China \cite{guo2010single}\.
Soybeans (Glycine max) primarily self-pollinate, outcrossing at a rate of 1-8% \cite{ray2003soybean}\.
Soybeans readily cross with their wild progenitor, G. soja \cite{singh1988genomic}\.
Plants of intermediate morphology appear around fields when S. soja is also present, indicating spontaneous hybridization \cite{kwon1972studies}\.
The accession "G. gracilis" shows allelic contribution from both G. max and G. soja \cite{keim1989restriction}\.

\subsection*{Palm}

Oil palm (Elaeis guineensis) of western Africa has been crossed with the New-World species E. oleifera, the only other species in the genus.
The two species differ in vegetative and reproductive morphology and in ecological adaptations.
The hybrid appears to have partial resistance to pathogenic fungal Phytophthora palmivora, native to Latin America \cite{torres2010first}\.
The F1 hybrids have moderately diminished viability and fertility, but the reproductive barrier is not fully developed \cite{hardon1969interspecific}\.
Since there is only one wild relative in the same genus, and has not been sympatric with E. guineensis for a long time it seems unlikely that adaptive introgression has played a significant role in the domestication of this plant.

\subsection*{Sugar Beet}

Sugar beet was domesticated relatively recently
Crop-wild weedy hybrids (wild alleles introgressing into crop populations) \cite{ellstrand2003dangerous}\
Most introgression into sugar beet from wild relatives has been the intentional product of breeding efforts for the purpose of disease resistance (a review of these efforts and outcomes can be found at \cite{panella2007broadening})\.

@@

\subsection*{Cotton}
Cited (indirectly) in \cite{westengen2005genetic}:
45 diploid species, 5 tetraploid
monophyletic
all tetraploids from New World
two domestications, G. hirsutum (Mesoamerica/Caribbean) and G. barbadense (S. America/Caribbean)

Cited (indirectly) in \cite{ellstrand1999gene}\.
Evidence of introgression from domesticated cottons into wild.

Directly found by \cite[wendel1989genetic}\:
Bidirectional interspecific introgression between cultivated and wild cottons account for a "significant" amount of allele diversity, in both domesticated species.
This is not shown to be adaptive, though.

\subsection*{Rapeseed}

Domesticated rapeseed (Brassica napus) has many sympatric wild relatives, but reproductive barriers limit the degree of hybridization that occurs.
\cite{scheffler1994opportunities}\ experimentally tested fertility rates between domesticated rapeseed and wild relatives.
Also see \cite{ellstrand1999gene}\.

\subsection*{Coconut}

\subsection*{Sorghum}

@@
